---
title: "Raman phenotyping"
author: "Ruben Props, Cristina Garcia Timermans and Frederiek-Maarten Kerckhof"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Raman Phenotyping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Raman phenotyping

This vignette illustrates how to extract microbial phenotypes out of microbial Raman data. It follows the same logic and strategy as the [Phenoflow package](https://github.com/CMET-UGent/Phenoflow_package).





## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
